Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques

### 2020-04-04 paper.tex
Move body of paper into texfiles/ and rename main document to paper.tex
Remove \EL{} macros

\definecolor{myred}{HTML}{b30000}
\definecolor{mygreen}{HTML}{008000}
\newcommand{\EL}[1]{{\color{myred}{#1}}}
\newcommand{\TL}[1]{{\color{mygreen}{#1}}}

\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}
\newcommand{\QQ}[1]{{\color{green}{#1}}}

\title{
\EL{Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques}
}

@@

\begin{abstract}
	\EL{This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events. 
	%
	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
	%
	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle. This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag. 
	%
	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.  These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
        The present study is the first application to a fluid-structure interaction problem. 
	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
	%
	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
	% 
	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.}
      This paper is supplemented by a software repository containing the codes for the simulations and data analysis involved in this study: \url{https://github.com/tlestang/paper_extreme_drag_fluctuations}.
\end{abstract}

\include{section1}
\include{section2}
\include{section3}
\include{section4}
\include{section5}

@@

\appendix
\include{appendixA}
\include{appendixB}
\include{appendixC}
\include{appendixD}

>>>>

\title{
Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques
}

@@

\begin{abstract}
	This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging on a bluff body, and examines the relevance of two distinct rare-event algorithms to efficiently sample these events. 
	%
	The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
	%
	Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle. This vortex is responsible for a significant pressure drop between the forebody and the base of the obstacle, thus yielding a very high value of the drag. 
	%
	Two algorithms are then considered to speed up the sampling of such flow scenarii, namely the \ac{ams} and the \ac{gktl} algorithms. 
	The general idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned, according to some specific rules designed to sample large-amplitude events more frequently.  These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry.
        The present study is the first application to a fluid-structure interaction problem. 
	% allowing the simulation of rare  events otherwise out of reach by direct sampling. 
	%
	Practical evidence is given that the fast sweeping time of turbulent fluid structures past the obstacle has a strong influence on the efficiency of the rare-event algorithm. 
	% 
	While the \ac{ams} algorithm does not yield significant run-time savings as compared to direct sampling, the \ac{gktl} algorithm appears to be effective to sample very efficiently extreme fluctuations of the time-averaged drag and estimate related statistics such as return times.
      This paper is supplemented by a software repository containing the codes for the simulations and data analysis involved in this study: \url{https://github.com/tlestang/paper_extreme_drag_fluctuations}.
\end{abstract}

\include{./texfiles/section1}
\include{./texfiles/section2}
\include{./texfiles/section3}
\include{./texfiles/section4}
\include{./texfiles/section5}

@@

\appendix
\include{./texfiles/appendixA}
\include{./texfiles/appendixB}
\include{./texfiles/appendixC}
\include{./texfiles/appendixD}

### 2020-04-04 texfiles/appendixA.tex
Remove \EL{} macros

% details about LBM
In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice \EL{(see \citep{kruger_lattice_2017} for a comprehensive introduction)}. 
In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
\[
\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
\]
for the mass density and the fluid momentum respectively. The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.

% algo
%
The complexity of the flow emerges from the repeated application of simple rules of streaming and collision. The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure. Namely, an \emph{exact} streaming step 
\[
f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
\]
during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
\[
f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
\]
which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level). The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by 
\[
\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
\]
This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \EL{\citep{BGK}}.
%
The equilibrium function is given by
\[
f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
\] 
with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e. $\Delta x \to 0$ \EL{\citep{succi_book}}.

>>>>

% details about LBM
In the LB method, the fluid is viewed as a population of particles that collide, redistribute and propagate along the different links of a discrete lattice (see \citep{kruger_lattice_2017} for a comprehensive introduction). 
In our two-dimensional situation, the so-called D2Q9 lattice with only nine possible velocities $\{\mathbf{c_i}\}_{i=0...8}$ at each node has been adopted (see  Fig.~\ref{fig:D2Q9}).
Locally, the macroscopic flow variables (per unit volume) are recovered by summing over the densities of particles $\{f_i\}_{i=0...8}$ moving with the different velocities, i.e.
\[
\rho(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \quad \mathrm{and}\quad \rho(\mathbf{x},t) \mathbf u(\mathbf{x},t) = \sum_i f_i(\mathbf{x},t) \mathbf{c_i}
\]
for the mass density and the fluid momentum respectively. The assumption of weak compressibility (for an ideal gas) is made so that the pressure is directly proportional to the mass density: $p = c_s^2 \rho$ where $c_s$ is interpreted as a speed of sound.

% algo
%
The complexity of the flow emerges from the repeated application of simple rules of streaming and collision. The \ac{lbm} advances the local densities of particles $f_i(\mathbf{x},t)$ moving with velocities $\mathbf{c}_i$  in a two-step procedure. Namely, an \emph{exact} streaming step 
\[
f_i(\mathbf{x}+\mathbf{c}_i \Delta t, t + \Delta t) = f_i^{\mathrm{out}}(\mathbf{x},t)
\]
during which particles move with their own velocity to a neighbouring node, and an instantaneous collision step
\[
f_i^{\mathrm{out}}(\mathbf{x},t) = -\frac 1 {\tau_\nu} \left(f_i(\mathbf{x},t) - f_i^\mathrm{eq}(\mathbf{x},t) \right)
\]
which achieves a relaxation of local densities towards an absolute equilibrium (at the macroscopic level). The time-scale $\tau_\nu$ (in lattice unit) is related to the kinematic viscosity of the fluid by 
\[
\nu = \left( {\tau_\nu} - \frac 1 2 \right) c_s^2 ~\Delta t
\]
This simplification of the collision kernel is known as the BGK approximation in the kinetic theory of gas \citep{BGK}.
%
The equilibrium function is given by
\[
f_i^\mathrm{eq}(\mathbf{x},t) = w_i  \rho(\mathbf{x},t) \left( 1 + \frac{\mathrm u(\mathbf{x},t) \cdot \mathbf{c_i}}{c_s^2} +
\frac{u_\alpha(\mathbf{x},t) u_\beta(\mathbf{x},t)({c_i}_\alpha {c_i}_\beta - c_s^2 \delta_{\alpha\beta})}{2 c_s^4} \right)
\] 
with the weight factors $w_0=4/9,~w_{1...4} = 1/9$ and $w_{5...8}=1/36$ for the D2Q9 lattice. 
This discrete Lattice Boltzmann scheme is second-order accurate in $\Delta x $ and compliant to the weakly-compressible Navier-Stokes equations with a third-order error in $\mathrm{Ma}=|\mathbf{u}|/c_s$ as the lattice spacing vanishes, i.e. $\Delta x \to 0$ \citep{succi_book}.

### 2020-04-04 texfiles/appendixB.tex
Remove \EL{} macros

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig21/fig21}
	\caption{\label{fig:illustr_AMS} \EL{Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$. $\xi(x(t))$ is a prescribed score function defined along each trajectory. 
	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$.} }
\end{figure}

The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
%
The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories. On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line). Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$. It is then simulated from this branching point to the final time $T_a$. In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories. This re-sampling procedure can \EL{be iterated} $J$ times or until all trajectories do exceed a fixed threshold $Q$.

@@

	\begin{figure}
		\centering
		\includegraphics[width=.7\linewidth]{fig22/fig22}
		\caption{Efficiency of the \EL{\ac{ams} algorithm} with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}. The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$. The blue line is the analytical solution for the return time of amplitude $a$.}
		\label{fig:comparaison_temps_de_retour}
	\end{figure}
	A one-dimensional \acl{ou} process is considered:
	\begin{equation}
	\label{eq:ou}
	\dot{x} = -x + \eta (t),
	\end{equation}
	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.

@@

	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents. Such favourable features \EL{do not} persist in the case of fluid dynamics.

>>>>

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{fig21/fig21}
	\caption{\label{fig:illustr_AMS} Operating principle of the \ac{ams} algorithm with three trajectories of duration $T_a$. $\xi(x(t))$ is a prescribed score function defined along each trajectory. 
	Selection and branching rules are applied in order to reach a maximum score value within the time interval $[0,T_a]$. }
\end{figure}

The operating principe of the AMS algorithm is sketched in Fig.~\ref{fig:illustr_AMS}.
%
The trajectories 1, 2 and 3 represent the evolution of the \emph{score function} for the current ensemble of trajectories. On the basis of their respective maximum: $Q_1$, $Q_2$ and $Q_3$, the trajectory with the lowest maximum is discarded in the ensemble (dashed line). Among the two remaining trajectories, trajectory 3 is chosen randomly and copied until it reaches the value $Q_1$. It is then simulated from this branching point to the final time $T_a$. In the present case of deterministic dynamics, a small perturbation is introduced at the branching to separate the trajectories. This re-sampling procedure can be iterated $J$ times or until all trajectories do exceed a fixed threshold $Q$.

@@

	\begin{figure}
		\centering
		\includegraphics[width=.7\linewidth]{fig22/fig22}
		\caption{Efficiency of the \ac{ams} algorithm with respect to direct sampling in the case of an Ornstein-Uhlenbeck process \citep{lestang_computing_2018}. The red line represents the evolution of the maximum obtained from re-sampled trajectories as a function of the computational cost $C_{AMS}$. The blue line is the analytical solution for the return time of amplitude $a$.}
		\label{fig:comparaison_temps_de_retour}
	\end{figure}
	A one-dimensional \acl{ou} process is considered:
	\begin{equation}
	\label{eq:ou}
	\dot{x} = -x + \eta (t),
	\end{equation}
	where $\eta$ is a Gaussian noise with $\langle \eta(t)\eta(t-t')\rangle = \delta(t-t')$.

@@

	The \acl{ou} process showcases the efficiency of the \ac{ams} algorithm.
	However, the state space is here one-dimensional and the choice of the score function is straightforward: it is $x$ itself.
	In addition, the noise term in Eq.~\eqref{eq:ou} has no correlation in time, which implies that newly generated trajectories quickly separate from their parents. Such favourable features do not persist in the case of fluid dynamics.

### 2020-04-04 texfiles/appendixC.tex
Remove \EL{} macros

 \EL{A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamics. 
%This is true for stochastic dynamics.
Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.   
	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.}
%
% 
%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones. We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
%
%
Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
%

>>>>

 A key feature of the \ac{gktl} algorithm is that all resampled trajectories are solutions of the original dynamics. 
%This is true for stochastic dynamics.
Nevertheless, it should be noted that a small random perturbation is introduced in the cloning procedure to force clones from the same trajectory to separate, \emph{i.e.} artificial randomness is introduced so that the cloning procedure is effective for deterministic dynamics.   
	As for the \ac{ams} algorithm, it has been checked that this perturbation did not affect the statistics of the sampled trajectories.
%
% 
%However, similarly to the case of \ac{ams}, randomness must be artificially introduced for the cloning procedure to be effective for deterministic dynamics in order to separate the clones. We have proceeded in the same way as as described in section~\ref{sec:ams_drag}.
%
%
Eventually, the sampled trajectories obtained with the \ac{gktl} algorithm can be used to compute the statistical properties of any observable with respect to the distribution $\mathbb{P}_{0}$ from the distribution $\mathbb{P}_{k}$ by using Eq.~\eqref{eq:Biased_Path_Approximation}.
%

### 2020-04-04 texfiles/appendixD.tex
Remove \EL{} macros

\begin{figure}
  \centering
\includegraphics[width=.7\linewidth]{fig23/fig23}
\caption{\EL{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}}
\label{fig:pdf_drag_with_perturbation}
\end{figure}

>>>>

\begin{figure}
  \centering
\includegraphics[width=.7\linewidth]{fig23/fig23}
\caption{\ac{pdf}s of (normalised) drag fluctuations obtained from the perturbed (at branching points) and unperturbed numerical simulations; $\epsilon$ is the relative amplitude of the perturbation.}
\label{fig:pdf_drag_with_perturbation}
\end{figure}

### 2020-04-04 texfiles/section1.tex
Remove \EL{} macros

\section{Introduction}

% general comments on physical problem %
%
Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \EL{\citep{kanev2010}}.
%

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these extreme events are rare. \EL{The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.}
%
%
{Rare-event sampling \EL{refers} to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
%	
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is \EL{conducted} using both very long time-series (direct sampling) and rare-event sampling techniques.

%
In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}. In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
%
However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%
%
%
Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
%
Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
%
%
\EL{Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%
Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.}
%
\EL{Another problem related to rare events is the estimation of parameters or statistics from a limited number of samples.  Several authors have proposed original strategies  \citep{Mohamad-ws,Blonigan-ws} such as sequential sampling. These approaches and rare event algorithms could also be considered for applications in wave-structure problems, ship motion and load acting on offshore platforms \citep{belenky-ws}.}
%
The paper is organized in two parts. The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow. This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
%
The motivation for this study is twofold.
Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations. This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them. This assessment is developed in the second part of the paper.

% Annonce du plan
The flow set-up is introduced and the dynamics related to typical
drag fluctuations is described in section~\ref{sec:test_flow}.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, \EL{the phenomenology of  extreme drag fluctuations} is investigated based on direct sampling.
Both the instantaneous drag and time-averaged drag are considered.
It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics. 
Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly. This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
As a specific \EL{successful} application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion end this work.

>>>>

buffer-name
\section{Introduction}

% general comments on physical problem %
%
Turbulent flows are important in a variety of natural phenomena, industrial and civil applications.
Their characteristic feature is the spontaneous development of intense and sporadic motions associated with extreme internal forces \citep{lesieur_book,donzis_sreenivasan_2010,Yeung}.
``Extreme'' refers here to fluctuations that can deviate from the mean value by ${\cal{O}}(10)$ standard deviations.
In engineering, the nature of such extreme dynamical events and their statistics are of crucial interest to predict excessive mechanical efforts, which can threaten the structural integrity of embedded structures \citep{kanev2010}.
%

From the viewpoint of chaotic dynamical systems, turbulence in fluids is linked to non-linearity and strong departure from statistical equilibrium \citep{KRAICHNAN}.
The use of analytical perturbative methods in identifying resonant interactions (among degrees of freedom) responsible for extreme fluctuations is unsuccessful.
Alternatively, simulation offers a practical approach to gain physical insight into these events, quantifying their intensity and estimating their frequency of occurrence.
However, this requires very long simulations since these extreme events are rare. The computational cost for sampling a fluctuation of very small probability typically grows as the inverse of this probability~\citep{wouters2016rare}.
%
%
{Rare-event sampling refers to a large body of methods that aim at \emph{preferentially} exploring  regions of phase space corresponding to  events that would otherwise be accessed with a very low probability through a brute-force direct sampling.}
%
%	
% next two paragraphs may be too technical -- Find a way to summarize them in a few sentences accessible to a broad audience in fluid mechanics? Notion of action will be vague for many readers
%
In the present work, a computational study of extreme mechanical efforts acting on an immersed bluff body is conducted using both very long time-series (direct sampling) and rare-event sampling techniques.

%
In fluid turbulence, rare-event sampling has been approached mainly from the perspective of simplified dynamics such as the one-dimensional Burgers' equation with a stochastic forcing \citep{bec_burgers_2007}. In this case, dynamics can be sampled by using a Markov chain Monte-Carlo algorithm \citep{duben_monte_2008,mesterhazy2011anomalous,mesterhazy2013lattice} that provides a framework for rare-event sampling.
%
An alternative approach is based on instantons \citep{gurarie_instantons_1996,grafke2015instanton} and applies to stochastically driven systems in the limit of weak noise.
Instantons refer to the most probable trajectories in phase space that achieve a given rare event (in the limit of weak noise). Suitable numerical schemes can be used to evaluate instantons as well as the related probabilities of rare events \citep{chernykh_large_2001,grafke_instanton_2013,grigorio_instantons_2017,laurie2015computation,bouchet2014langevin}.
An example is the investigation of the physics of rogue waves~\citep{dematteis2018rogue,dematteis2019experimental}.
%
However, a drawback of the aforementioned approaches is their limitation to simple and stochastically driven dynamics.
%
%
%
Here, a more general approach is considered for complex, possibly deterministic, dynamical systems.
It is based on sampling algorithms relying on \emph{selection rules} applied to an ensemble of trajectories of the system. 
%
Even though such ideas date back to the early 1950s, they have received growing interest over the last twenty years with successful applications in various domains such as chemistry \citep{van_erp_elaborating_2005,escobedo_transition_2009,teo_adaptive_2016}, biophysics \citep{huber_weighted-ensemble_1996,zuckerman2017weighted,bolhuis2005kinetic}, nuclear physics \citep{louvin2017}, nonlinear dynamical systems \citep{tailleur_probing_2007} and communication networks simulation \citep{villen-altamirano_restart:_1994}.
More importantly, such algorithms have been shown to be useful for the study of rare events in simple deterministic dynamics~\citep{wouters2016rare}.
%
%
Certainly, an original contribution of our study is to test the application of rare-event sampling algorithms in the context of far-from-equilibrium dynamics with an irreducible very large number of degrees of freedom. 
%
Two algorithms that are a priori suitable for such dynamics are considered, namely, the Adaptive Multilevel Splitting (AMS) algorithm \citep{cerou_adaptive_2007} and the Giardina-Kurchan-Tailleur-Lecomte (GKTL) algorithm \citep{giardina_direct_2006}.
%
Another problem related to rare events is the estimation of parameters or statistics from a limited number of samples.  Several authors have proposed original strategies  \citep{Mohamad-ws,Blonigan-ws} such as sequential sampling. These approaches and rare event algorithms could also be considered for applications in wave-structure problems, ship motion and load acting on offshore platforms \citep{belenky-ws}.
%
The paper is organized in two parts. The first part highlights the phenomenology of extreme fluctuations of the drag force acting on a square placed in a two-dimensional turbulent channel flow. This study is based on the simulation of the flow over a very long duration, made possible by the relative simplicity of the flow.
%
The motivation for this study is twofold.
Firstly, it provides a detailed description of the statistics and dynamics related to extreme drag fluctuations. This analysis is informative from the  viewpoint of fluid mechanics and, to the best of our knowledge, has never been reported before.
Secondly, it yields reference results that are required to validate the outputs of rare-event algorithms and to evaluate the possible computational gain obtained from them. This assessment is developed in the second part of the paper.

% Annonce du plan
The flow set-up is introduced and the dynamics related to typical
drag fluctuations is described in section~\ref{sec:test_flow}.
The statistical properties of the drag are then discussed.
In section~\ref{sec:direct_sampling}, the phenomenology of  extreme drag fluctuations is investigated based on direct sampling.
Both the instantaneous drag and time-averaged drag are considered.
It is found in particular that sampled extreme fluctuations of the instantaneous drag result from very similar dynamics. 
Section~\ref{sec:rare_events_algorithms} examines the applicability of both the \ac{ams} and \ac{gktl} algorithms to the simulation of extreme drag fluctuations in the same flow configuration.
In subsection~\ref{sec:ams}, we show that the use of the \ac{ams} algorithm is not successful, or at least not straightforwardly. This difficulty is put in perspective with the phenomenology of extreme drag fluctuations developed in  section \ref{sec:direct_sampling}.
Subsection~\ref{sec:gktl} presents the computation of extremes of the time-averaged drag by using the \ac{gktl} algorithm.
This latter allows for an exceptional reduction of the computational cost needed to simulate trajectories corresponding to extreme values of the time-averaged drag.
As a specific successful application, the \ac{gktl} algorithm is used to compute the return times of extreme fluctuations of the time-averaged drag acting on the immersed obstacle.
Perspectives and conclusion end this work.
